Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest

### 2020-01-23 Manuscript.Rmd
Uppangala 0.1.9057 23.20 MRM & VR review

---
title: Environment filter different tree functional composition between canopy and understory guilds in India / Different environment filters shape functional composition of canopy and understory guilds in Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - email
    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

@@

> The abstract is just a first try, maybe ignore it for the moment until the introduction and the discussion is written.

1. Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities. 
<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.  -->
Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests.
We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.

1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
The plot was subdivided into tree communities, each composed of guilds of canopy and understory tree.
We combined tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits to assess the variation of functional composition of between understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.

1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
Although functional traits did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. 
Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.

1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences tree community composition at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Specifically, topography can filter different ecological strategies [@kraft_functional_2008] due to variation in water and nutrient availability in tropical forests [@ferry2010higher; @Blanchard2019] .
<!-- + Blanchard, Grégoire, François Munoz, Thomas Ibanez, Vanessa Hequet, Hervé Vandrot, Jérémy Girardi, et Philippe Birnbaum. 2019. « Regional rainfall and local topography jointly drive tree community assembly in lowland tropical forests of New Caledonia ». Journal of Vegetation Science 0 (ja). https://doi.org/10.1111/jvs.12781. -->
At higher topographic positions, e.g. hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review].

In addition, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Stark2015; @Uriarte2004a]. Tree species can occupy different vertical strata at adult stage, allowing to classify them into vertical guilds. Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical stratification is little addressed in tropical forests [@wright2002plant].

<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011]. Therefore, understory and canopy guilds can differently respond to environmental gradients. Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests [@wright2002plant].  Whereas, investigating community assembly among vertical strata may reveal distinct response to environmental gradients [@Raevel2018], and thus provide a better understanding of community assembly processes. -->

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes. Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. 
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species.
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].

Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds. Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018]. In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), while limited light availability should favor conservative strategies in understorey trees (along the leaf economics spectrum). Thus we expected different trait-environment relationships across vertical guilds, depending on the tree species strategies along the fundamental leaf and wood ecnonomics spectra. Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013], but evidences are lacking for Asian tropical forests.

We assessed the variation in functional composition of understorey and canoy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations. We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000]. The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play a role.

We used tree census data, LiDAR-derived topographic data and canopy metrics, and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do the mean trait variations among tree species follow independent leaf and wood economics spectra at local scale ? If so, how are canopy and understory guilds distributed along leaf and wood economics spectra ? (ii) Does functional composition differ between assemblages of canopy and understory guilds ? and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently ?

<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species. We hypothesized functional composition to differ between canopy and understorey species, due to their differences in growth and functional strategies [@Herault2011]. Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.  -->

@@

The study was carried out in a nearly 10-ha (300m*330m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996]. The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996]. The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

@@

All trees $\geq$ 10 cm in diameter at breast height (dbh) have been identified to species level, tagged, mapped and monitored since 1989 following international standards [@Jairazbhoy2006], in an area progressively extended to a contiguous 10-ha plot in 2014.
The plot includes 6672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39m, canopy between 25 and 39m, sub-canopy between 15 and 25m, and understory below 15m. Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot. We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80.1%) species, representing 99.0% of tree individuals in the plot. We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species, depending on leaf accessibility and species abundance (3 individuals on average). For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter close to 1 cm, following recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched CO2 air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (1 cm in diameter) to allow measuring branch wood density. Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively.

@@

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$). Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an analytical balance (Wensar, KD-H600, 0.01g), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of [@Perez-Harguindeguy2013].

@@

We used LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure. The LiDAR campaign was conducted in 2013, consistent with field measurements. Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher]. In addition, based on a 1-m resolution LiDAR model of canopy (Fig. \@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.

@@

To assess the degree of trait variation and covariation we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.

We subsampled tree communities in $30m*30m$ subplots (n=110), and characterized the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.

We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’, [@Raevel2018]) for each trait in each subplot:

@@

We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild, [@bernard-verdier_community_2012]). The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances. We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as

$$SES = \frac{GWM_{obs}-\overline{GWM_null}}{\sigma_{null}}$$
where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.

We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions and a bidirectional stepwise selection of predictors. These analyses were performed independently for each of the five functional traits. All analyses were conducted using the R 3.3.0 software [@RCoreTeam2018].

@@

The first axis of the PCA performed on species trait values (Fig. \@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC. WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information. Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3, respectively Fig. \@ref(fig:pca)B).

@@

At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig. \@ref(fig:GWMboxplot)). GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, higher for LA and lower for LDMC and LT than understory tree assemblages. In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.

We detected a significant effect of slope, curvature and wetness on both canopy and understory guilds assemblages, of southwesterness on canopy guilds assemblages and of canopy height on understory guild assemblages (Table \@ref(tab:LM)).

> *Un peu court/abrupt. Préciser les directions des effets observés.* François

>>>>

---
title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - email
    - Forestry and Ecology Group, National Remote Sensing Centre, Hyderabad 500 037, India
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

@@

1. Topographic heterogeneity determines changing soil conditions, and environment also changes across vertical strata in rainforests. However, the joint influence of these factors on tree functional composition remain little known.
<!-- Joint study of topography and community vertical structure hold promise to understand the drivers of hyperdiverse tropical communities.  -->
<!-- Moreover, understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition.  -->
<!-- Despite, assembly of tree communities has been mostly studied irrespectively of vertical stratification in tropical forests. -->
We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity, depending on resource availability, hydraulic and biomechanic constraints.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. 
The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees.
We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.  
1. The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra, even at local scale. 
Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. 
Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities.
Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

<!-- > To be checked at the end -->
Rainforest; Western Ghats; Economics spectra; Environmental filtering, Vertical strata

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003]  and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Specifically, topography is known to filter different ecological strategies [@kraft_functional_2008] due to its impact on water and nutrient availability [@ferry2010higher; @Blanchard2019].
<!-- At higher topographic positions, e.g., hilltops, decreased water and nutrient availability  [@ferry2010higher] is expected to filter species with more conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review]. -->
However, tree species can occupy different vertical strata at adult stage, and thus be classified into vertical guilds (understory and canopy), experimenting contrasted environmental conditions. 
Indeed, light availability decreases downward across vertical forest strata, while hydraulic and mechanical constraints increase upward, so that different survival and growth strategies should be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a].
While the influence of ground-level variation in environmental conditions (soil, topography) on community assembly is now widely recognized, environmental filtering across vertical guilds is little addressed in tropical forests [@wright2002plant].

<!-- In addition, different responses of vertical strata to environmental gradients have been observed in Columbia [@duque2002different] and central Africa [@Ouedraogo2013]. -->

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], contributing thus to test for community assembly processes. Functional traits have been defined as phenotypic traits that impact fitness indirectly through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. 
Understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].
However, functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. 
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients.
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012].
Nevertheless, @reich_world-wide_2014 proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].

<!-- Here we decompose a tree community in spatially circumscribed at ground level into vertical guilds. Because they experience different environmental constraints, vertical guilds should show different functional responses to environmental variation between communities [@Raevel2018].  -->
In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; Schmitt et al., under review] of canopy trees to water and nutrient limitation on hilltops [@ferry2010higher], 
 <!-- In the context of seasonal rainforests, we hypothesized that greater hydraulic constraints should entail stronger response of canopy trees to water limitation on hilltops (along the wood economics spectrum), -->
while limited light availability should favor conservative strategies in understory trees (along the leaf economics spectrum). Thus we expected different trait-environment relationships across vertical guilds, depending on tree species strategies along the fundamental leaf and wood economics spectra. Previous works have shown different response of vertical strata to environmental variation at ground level, in Columbia [@duque2002different] and central Africa [@Ouedraogo2013].

We assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, and examined the effects of abiotic and biotic environments on these variations. We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000]. The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, we expected hydraulic and mechanic constraints to play an important role in tree community assembly.

We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, in multivariate approaches and linear regressions to address the following questions: (i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? (ii) Does functional composition differ between assemblages of canopy and understory guilds? and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?

<!-- Based on previous evidences of independent economics spectra at large scale in the Neotropics [@baraloto_decoupled_2010; @fortunel_leaf_2012], we would expect leaf and wood economics spectra to be maintained in Indian rainforests at local scale [but see @Messier2017], with differentiated trait spaces for canopy and understory species. We hypothesized functional composition to differ between canopy and understory species, due to their differences in growth and functional strategies [@Herault2011]. Because this region experiences a long dry season period and is under the influence of intense monsoon rainy seasons with strong wind disturbances, we expected that the local topography is a major driver of tree community functional composition.  -->

@@

The study was carried out in a nearly 10-ha (300 m*330 m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996]. The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall of 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996]. The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

@@

All trees $\geq$ 10 cm in diameter at breast height (dbh) have been identified to species level, tagged, mapped and monitored since 1989 following international standards [@Jairazbhoy2006], in an area progressively extended to a contiguous 10-ha plot in 2014.
The plot includes 6,672 tree individuals belonging to 111 species mostly belonging to the Dipterocarpaceae, Myristicaceae and Fabaceae families.

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species (vertical stratum reached by most adult trees, with four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. Because our study site is characterized by a steep topography, sub-canopy species can often reach the canopy in the Uppangala plot. We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density. 
<!-- Among the selected species, 54 and 35 belonged to the canopy and understory guilds, respectively. -->

@@

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$). Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an analytical balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.

@@

We used airborne LiDAR (Light Detection And Ranging) data to characterize the local topography and forest canopy structure. The LiDAR campaign was conducted in 2013, consistent with field measurements. Using a 1-m resolution LiDAR digital elevation model, we derived four statistically independent (Pearson’s r < 0.48) topographical variables: (i) curvature, related to local water and soil accumulation areas; (ii) wetness, an index accounting for both slope and elevation and highlighting water flow paths and accumulation at the study area scale [@Bohner2006], (iii) southwesterness (SW), representing the topographical aspect in the direction of south-west: $cos(aspect~in~degree+225))$ to account for the exposure to dominant monsoon winds [@Bunyan2015]; and (iv) slope, which could have both a strong effect on forest dynamics and on nutrient/water availability [@ferry2010higher]. In addition, based on a 1-m resolution LiDAR model of canopy (Fig. \@ref(fig:Uppangala)), we calculated the top canopy height, as a proxy of decreasing light availability downward and thus of local competition for light.

@@

To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.

We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds.

We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:

@@

We tested whether variation in functional composition depended on local topography and canopy height for each guild, based on a null model randomizing trait values among species within each guild (permuting rows in Trait*Species matrices of each guild; [@bernard-verdier_community_2012]). The null model broke the link between environment and local functional composition while maintaining the spatial distribution of guilds and species abundances. We derived 999 randomized GWM values and calculated the standardized effect sizes (SES) quantifying the deviation of observed values from the null distribution as

$$SES = \frac{GWM_{obs}-\overline{GWM_{null}}}{\sigma_{null}}$$
where $GWM_{obs}$ was the observed GWM, and $\overline{GWM_{null}}$ and $\sigma_{null}$ were the mean and standard deviation of the null distribution, respectively.

We assessed how the deviation of guild composition from the null model depended on environmental variables by performing linear regressions and a bidirectional stepwise selection of predictors. These analyses were performed independently for each of the five functional traits. All analyses were conducted using the R 3.3.0 software [@RCoreTeam2020].

@@

The first axis of the PCA performed on species trait values (Fig. \@ref(fig:pca)A) was mostly related to wood density and structural leaf traits (LT, LA, LDMC), and opposed species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC. WD was orthogonal to SLA in the plan 1-2 of the PCA suggesting that WD and SLA provide independent information. Canopy and understory species did not show significant differentiation along the first and second PCA axis (Wilcoxon p=0.2 and p=0.3 respectively, Fig. \@ref(fig:pca)B).

@@

At community level, the guild weighted mean (GWM) of all traits significantly differed between understory and canopy guilds (Student p<0.001, Fig. \@ref(fig:GWMboxplot)). GWM values of canopy assemblages were lower for SLA and WD with almost no overlap and, to a lesser extent, lower for LA and higher for LDMC and LT than understory tree assemblages. In other words, understory guild assemblages had larger, thinner and lighter leaves with higher SLA and denser woods, compared to canopy guild assemblages.
